RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-10-08 formalframework.tex
updated into formal framework

In Section~\ref{sect:gmas} we developed a list of critical questions and argument schemes by analyzing transcripts of discussions about the development of a traffic simulator. The resulting list is shown in Table~\ref{table:argument-schemes}. We then presented the RationalGRL framework in secton~\ref{sect:overview}, consisting of a modeling language, a metamodel, and various examples of using the argument schemes and critical questions. In section~\ref{sect:methodology} we explained how this framework can be used by practitioners by explaining the methodology and tool support.

@@

In the first two subsections we formalize a static representation of our framework: We provide a formal specification of a GRL model based on the GRL metamodel (Section \ref{sect:goalmodeling:logicallanguage}) in the first subsection, and we extend this with arguments and attack links in the second subsection, hereby obtaining a formal specification of a RationalGRL model. In the third subsection we turn to the dynamics of our framework. We develop algorithms for instantiating argument schemes and for answering critical questions. We show that each algorithm is correct, meaning that if the RationalGRL model was valid before applying the algorithm, it is also valid afterwards.

@@

We start with the definition of various sets of elements that we use in subsequent definitions.

\begin{definition}[Set definitions]
\label{def:set-definitions}
Throughout this section, we adopt the convention that variables start with a lowercase letter, and sets and constants start with an uppercase.
\label{def:ie}
We define the following sets:
\begin{itemize}
\item $Type = \{Softgoal, Goal, Task, Resource\}$,
\item $Names$ is a finite set of string.
\item $DecompType = \{AND,OR,XOR\}$.
\item $ContribValue = \{Break, Hurt, Some\ nega-$ $tive, Unknown, Some\  positive, Help, Make\},$
\end{itemize}
\end{definition}

@@

>>>>

In Section~\ref{sect:gmas} we developed a list of critical questions and argument schemes by analyzing transcripts of discussions about the development of a traffic simulator. The resulting list is shown in Table~\ref{table:argument-schemes}. We then presented the RationalGRL framework in secton~\ref{sect:overview}, consisting of a modeling language, a metamodel, and various examples of using the argument schemes and critical questions. In section~\ref{sect:meth+tool} we explained how this framework can be used by practitioners by explaining the methodology and tool support.

@@

In the first two subsections we formalize a static representation of our framework: We provide a formal specification of a GRL model based on the GRL metamodel (Section \ref{sect:goalmodeling:logicallanguage}) in the first subsection, and we extend this with arguments and attack links in the second subsection, hereby obtaining a formal specification of a RationalGRL model. We then provide a formal translation procedure from a RationalGRL model to a GRL model.

In the third subsection we turn to the dynamics of our framework. We develop algorithms for instantiating argument schemes and for answering critical questions.

The two main results of this section are that (i) the translation procedure from a RationalGRL model to a GRL model is correct, meaning that if the RationalGRL model is valid, the resulting GRL model is also valid, (ii) the algorithms for applying argument schemes and critical questions are correct as well, meaning that if the Rational model is valid before, it is also valid afterwards. These two results together entail that our framework and the corresponding operations on it are consistent.

@@

We start with some general definition that we use in subsequent definitions.

\begin{definition}[General definitions]
\label{def:set-definitions}
Throughout this section, we adopt the convention that variables start with a lowercase letter (e.g, $id, i, j, name, ie, goal$, and sets and constants start with an uppercase letter (e.g., $Type, AND, Goal$).

We define the following sets:
\begin{itemize}
\item $Type = \{Softgoal, Goal, Task, Resource\}$,
\item $Names$ is a finite set of string.
\item $DecompType = \{AND,OR,XOR\}$.
\item $ContribValue = \{Break, Hurt, Some\ nega-$ $tive, Unknown, Some\  positive, Help, Make\},$
\end{itemize}
\end{definition}

@@

\paragraph{Example} \todo{Marc}{Marc}{Give example here}
